Spectroscopic confirmation of very-wide binaries and large-separation comoving stars from \textsl{Gaia DR1}

### 2017-07-11 paper/comoving-followup.tex
smoh comments

% Project-specific
\newcommand{\gaia}{\project{Gaia}}
\newcommand{\DR}[1]{\acronym{DR}#1}
\newcommand{\tgas}{\acronym{TGAS}}
\newcommand{\lbfgsb}{\texttt{L-BFGS-B}}
\newcommand{\Ha}{\ensuremath{{\rm H}\alpha}}
\newcommand{\dex}{\ensuremath{{\rm dex}}}
\newcommand{\llrold}{\ensuremath{\mathcal{R}_\mu}}
\newcommand{\llrnew}{\ensuremath{\mathcal{R}_{\rm RV}}}

@@

\begin{abstract}
% Context
Widely-separated ($\gtrsim X$~pc), coeval stars are powerful tools for testing star-formation
model predictions, improving stellar atmosphere models, and inferring the
mass distribution of the Milky Way.
\smoh{The following sentence seems to be in the air while I get what you mean.}
They vastly outnumber the Galactic open clusters, but are still expected to
share a common origin and chemistry.
\smoh{How about sticking with the terminology ("comoving (star) pair") we want settled?}
The kinematic properties of comoving pairs including bound wide binaries
serve as potentiometers for the smooth tidal field and perturbation spectrum of
the Milky Way.
\smoh{If you are referring to Shaya \& Olling, how about adding 'using HIPPARCOS'?}
Thus far, a few hundred genuine comoving star pairs with separations as large
as $8~\pc$ have been discovered in the solar neighborhood.
% Aims
In previous work, we have identified $\sim 10,000$ candidate comoving star pairs
using only astrometric data from the Tycho-Gaia Astrometric Solution (TGAS)
catalog, a part of \gaia\ \DR{1}.
Without radial velocity measurements for the vast majority of the pairs, the
\smoh{I prefer "false-positive rate" to "contamination rate" because it conveys
the meaning more directly.}
contamination rate at large separations is significant ($\sim 50\%$).
% Methods
\smoh{I think some details of telescope/instrument or just R~2000 should be mentioned here.}
Here, we present results from our own low-resolution radial-velocity survey of a sample of the
candidate pairs to validate wide-separation comoving star pairs.
% Results
Of the 311 observed comoving pairs, we identify 128 genuine comoving pairs with
separations as large as $\approx 10~\pc$, our original search limit.
% Conclusions
We find that the number of comoving star pairs with very wide separations
appears flat in linear separation at separations $\gtrsim 1~\pc$, strongly
suggesting the existence of a population of recently unbound wide binaries or
stellar associations.
% Gaia DR2 -> how many do we expect? Can we say anything? Out to what distance and
% separation for a typical MS star can we determine they are bound assuming
% high-res spectra to get very precise RVs?
\end{abstract}

@@

The feasibility of chemical tagging depends on not only on the precision of
the abundance measurements but also on the intrinsic chemical
dispersions of stars formed from the same birth clusters,
and the ability to tell that the chemical abundances of
two stars from the same cluster are more similar than two randomly-paired Galactic disk stars.
Recent works suggest that, at least for red giant stars, the abundance
distributions of 20 elements in a handful of open clusters included in the
\acronym{APOGEE} survey (\citealt{Majewski:2016,Nidever:2015}) are consistent
with having intrinsic dispersions smaller than the abundance measurement
precisions ($\approx 0.03~\dex$; \citealt{Bovy:2016,Ness:2017}).
\smoh{I find the following sentence out of place. Doesn't it mean that
  the intrinsic dispersion is effectively zero rather than ``unconstrained''?}
In effect, for most elements, the intrinsic dispersions are unconstrained
because of measurement noise, so the true intrinsic dispersions are not known.
This demonstration of chemical uniformity in open clusters is promising for
chemical tagging.
However, it was also recently shown that $\approx 0.3\%$ of random pairings of
Galactic field red giant stars have abundance patterns that are
indistinguishable at the same level (``doppelgangers''), even with 20 chemical
species (\citealt{Ness:2017}).

@@

In previous work, we have identified a sample of high-confidence, candidate
comoving star pairs (\citealt{Oh:2017}) using only astrometric information from
the \tgas\ catalog (\citealt{Michalik:2015,Gaia-Collaboration:2016a}).
These comoving pairs were selected using a cut on a marginalized likelihood
ratio computed for all nearly co-spatial pairs of stars.
Briefly, likelihood 1, $L_1$, considers a model in which the full-space
velocities of each star in the pair are identical and sampled from a disk-like,
Sun-relative velocity distribution, and likelihood 2, $L_2$, considers a model
in which the full-space velocities of each star were drawn independently from
the same velocity distribution used above. The computed likelihood ratio is
actually the ratio of fully marginalized likelihoods,
$\mathcal{L}_1/\mathcal{L}_2$, after marginalizing over true velocity and
distance for each star in a given pair.
After a parallax signal-to-noise ratio (SNR) cut ($\pi/\sigma_\pi > 8$), we
compute the likelihood ratio for all stars within $10~\pc$ of each target star
from the \tgas\ catalog.
\smoh{To be complete, we should mention $\Delta v_\perp < 10$~km/s cut.
  Or I think it's also an option to cut on much of these details consistently.}
The likelihoods appropriately take into account the reported covariances and
uncertainties between the astrometric measurements in the \tgas\ catalog and
naturally handle geometric or projection effects in comparing spherical \smoh{?}
velocity components for star pairs separated by large angles on the sky.
After applying a conservative cut on the likelihood ratio motivated by computing
the same likelihood ratio for random pairings of stars ($\ln
\mathcal{L}_1/\mathcal{L}_2 > 6$, we find 13,058 comoving star pairs with
physical separations between $10^3~\au$ and $10~\pc$.
Of these pairs, there are 10,606 unique sources, implying that a significant
fraction of the identified pairs are members of larger groups.
$\approx$4,000 of the pairs are mutually exclusive, i.e. neither member is
paired with other stars.
For a full description of the selection method, see \citealt{Oh:2017}.

@@

We observed a total of 765 unique sources over 5 nights from (UT) 2017-03-11 to
2017-03-15, corresponding to 323 unique star pairs and $\approx 120$
calibration targets.
\smoh{How about ``The calibration targets serve two purposes. First, ...
  just so that it is more explicit, plain, and direct, and also ties in
  with how we chose the calibration targets?}
The calibration targets have previously measured radial velocities and chemical
abundances, which will be used for validation of our reduction pipeline and in
future work to measure metallicities and stellar parameters for each observed
source using \project{The Cannon} (\citealt{Ness:2015}).
The spectra were reduced and calibrated using a custom, publicly-available
pipeline written in \python, described below.
\smoh{How the calibration targets were chosen is missing here.}

\smoh{Throughout this paragraph, consider changing e.g.,
  ``16 pixel rows'' to ``16 pixels in wavelength direction''?
  Or explicitly mention somewhere (perhaps at the start of this section)
  that x (along rows) is spatial and y (along columns) is spectral direction?}
For each source or comparison lamp observation, the two-dimensional image is
bias corrected, flat-field corrected, and trimmed using standard routines
implemented in \package{ccdproc} (\citealt{Craig:2015}).
For comparison lamp observations, the central 100 pixel columns are extracted
and summed along the spatial axis, weighted by the inverse-variances of each
pixel.
Source spectral traces were always placed in this region during observations,
and the curvature of the wavelength solution over this region is negligible.
For source observations, 1D spectra are extracted using an empirical
estimate of the pixel-convolved line spread function (LSF): at each of 16 pixel
rows between indices 750 and 850 (i.e. around the effective center of the
dispersion axis), a model for the LSF is fit to the 1D trace.
The LSF is assumed to be a Voigt profile with parameters $A$, amplitude; $x_0$,
profile centroid; $\sigma_G$, Gaussian root-variance; $\gamma_L$, Lorentzian
half-width at half maximum, and the background is modeled using a linear
polynomial with parameters $\alpha_1$ and $\alpha_2$, the slope and intercept
at the profile centroid, respectively.
The maximum-likelihood LSF + background model parameters are estimated from
each pixel row using \lbfgsb\ optimization (\citealt{Zhu:1994}), implemented in
the \package{scipy} package.
From the 16 fits, the median profile width parameters, $\sigma_G$ and
$\gamma_L$, are then taken to be the LSF width parameters.
At each of 1600 rows in each source spectrum image, the LSF amplitude and
centroid, and background slope and intercept at line centroid, are then fit
using the same procedure as above but with the width parameters fixed.
The LSF amplitudes are stored as the source fluxes, and the background model
intercepts are stored as the background (sky) fluxes.
Occasionally, nearby sources fall in the slit; these are masked by hand in the
2D images and do not bias the LSF extraction fits.

@@

This wavelength calibration procedure produces a nightly model for the
nonlinear dispersion of wavelength as a function of pixel value for each
extracted 1D spectrum.
However, because of flexure in the spectrograph, each observation will have an
additional offset of, typically, $\approx 0.5$--$2~{\rm \AA}$, corresponding to
a significant velocity offset at \Ha\ of $\approx 25$--$100~\kms$ (discussed
below).
\figurename~\ref{fig:sample-spectra} shows a random sample of 10
wavelength-calibrated source spectra of comoving star pair members normalized
to 1 at $\lambda = 5500~{\rm \AA}$ and \smoh{shifted by index (confusing)} for visualization.
As suggested by the features in these 10 spectra, there is a mix of
metallicities and spectral types amongst the observed targets.
Note that the spectra are not flux-calibrated, as we only care about measuring
radial velocities from this sample.

>>>>

% Project-specific
\newcommand{\gaia}{\project{Gaia}}
\newcommand{\DR}[1]{\acronym{DR}#1}
\newcommand{\tgas}{\acronym{TGAS}}
\newcommand{\lbfgsb}{\texttt{L-BFGS-B}}
\newcommand{\Ha}{\ensuremath{{\rm H}\alpha}}
\newcommand{\dex}{\ensuremath{{\rm dex}}}
\newcommand{\llrold}{\ensuremath{\mathcal{R}_\mu}}
\newcommand{\llrnew}{\ensuremath{\mathcal{R}_{\rm RV}}}
\newcommand{\npairsobs}{311}
\newcommand{\ncomoving}{127}

@@

\begin{abstract}
% Context
Widely-separated ($\gtrsim 0.01$~pc), coeval stars are powerful tools for
testing star-formation model predictions, improving stellar atmosphere models,
and inferring the mass distribution of the Milky Way.
They likely outnumber the Galactic open clusters by a large margin, but are
still expected to share a common origin and chemistry.
The kinematic properties of comoving pairs---including bound wide binaries---
serve as potentiometers for the smooth tidal field and mass perturbation
spectrum of the Milky Way.
Thus far, a few hundred genuine comoving star pairs with separations as large
as $8~\pc$ have been discovered in the solar neighborhood using HIPPARCOS.
% Aims
In previous work, we have identified $\sim 10,000$ candidate comoving star pairs
using only astrometric data from the Tycho-Gaia Astrometric Solution (TGAS)
catalog, a part of \gaia\ \DR{1}.
Without radial velocity measurements for the vast majority of the pairs, the
false-positive rate at large separations is significant ($\sim 50\%$).
% Methods
Here, we present results from our own low-resolution radial-velocity survey of a
sample of the candidate pairs to validate wide-separation comoving star pairs.
% Results
Of the \npairsobs\ observed comoving pairs, we confirm \ncomoving\ comoving
pairs with separations as large as $\approx 10~\pc$, our original search limit.
% Conclusions
We find that the number of comoving star pairs with very wide separations
appears flat in linear separation at separations $\gtrsim 1~\pc$, strongly
suggesting the existence of a population of recently unbound wide binaries or
stellar associations.
% Gaia DR2 -> how many do we expect? Can we say anything? Out to what distance and
% separation for a typical MS star can we determine they are bound assuming
% high-res spectra to get very precise RVs?
\end{abstract}

@@

The feasibility of chemical tagging depends on not only on the precision of
the abundance measurements but also on the intrinsic chemical
dispersions of stars formed from the same birth clusters,
and the ability to tell that the chemical abundances of
two stars from the same cluster are more similar than two randomly-paired Galactic disk stars.
Recent works suggest that, at least for red giant stars, the abundance
distributions of 20 elements in a handful of open clusters included in the
\acronym{APOGEE} survey (\citealt{Majewski:2016,Nidever:2015}) are consistent
with having intrinsic dispersions smaller than the abundance measurement
precisions ($\approx 0.03~\dex$; \citealt{Bovy:2016,Ness:2017}), though the true
intrinsic dispersions are not known.
This demonstration of chemical uniformity in open clusters is promising for
chemical tagging.
However, it was also recently shown that $\approx 0.3\%$ of random pairings of
Galactic field red giant stars have abundance patterns that are
indistinguishable at the same level (``doppelgangers''), even with 20 chemical
species (\citealt{Ness:2017}).

@@

In previous work, we have identified a sample of high-confidence, candidate
comoving star pairs (\citealt{Oh:2017}) using only astrometric information from
the \tgas\ catalog (\citealt{Michalik:2015,Gaia-Collaboration:2016a}).
These comoving pairs were selected using a cut on a marginalized likelihood
ratio computed for all nearly co-spatial pairs of stars.
Briefly, likelihood 1, $L_1$, considers a model in which the full-space
velocities of each star in the pair are identical and sampled from a disk-like,
Sun-relative velocity distribution, and likelihood 2, $L_2$, considers a model
in which the full-space velocities of each star were drawn independently from
the same velocity distribution used above. The computed likelihood ratio is
actually the ratio of fully marginalized likelihoods,
$\mathcal{L}_1/\mathcal{L}_2$, after marginalizing over true velocity and
distance for each star in a given pair.
After a parallax signal-to-noise ratio (SNR) cut, $\varpi/\sigma_\varpi > 8$),
we compute the likelihood ratio for all stars within $10~\pc$ and with a
difference in tangential velocity $\Delta v_\perp < 10~\kms$ of each target star
from the \tgas\ catalog.
The likelihoods appropriately take into account the reported covariances and
uncertainties between the astrometric measurements in the \tgas\ catalog and
naturally handle geometric or projection effects in comparing spherical
velocities for star pairs separated by large angles on the sky.
After applying a conservative cut on the likelihood ratio motivated by computing
the same likelihood ratio for random pairings of stars ($\ln
\mathcal{L}_1/\mathcal{L}_2 > 6$, we find 13,058 comoving star pairs with
physical separations between $10^3~\au$ and $10~\pc$.
Of these pairs, there are 10,606 unique sources, implying that a significant
fraction of the identified pairs are members of larger groups.
$\approx$4,000 of the pairs are mutually exclusive, i.e. neither member is
paired with other stars.
For a full description of the selection method, see \citealt{Oh:2017}.

@@

We observed a total of 765 unique sources over 5 nights from (UT) 2017-03-11 to
2017-03-15, corresponding to 323 unique star pairs and $\approx 120$
calibration targets.
For the purposes of this work, all of the calibration targets have previously measured radial velocities and were selected from high-resolution spectroscopic
surveys \citealt{Bensby:2014,Pinsonneault:2014}; these stars are used to
validate our radial velocity measurements.
These stars also have measured chemical abundances, and will be used in future
work as a training set for \project{The Cannon} (\citealt{Ness:2015}) to infer
metallicities and stellar parameters for our observed sources.
The spectra were reduced and calibrated using a custom, publicly-available
pipeline written in \python, described below.

For each source or comparison lamp observation, the two-dimensional image is
bias corrected, flat-field corrected, and trimmed using standard routines
implemented in \package{ccdproc} (\citealt{Craig:2015}).
The 2D spectra are oriented such that the spatial direction is along a row of
pixels, and the dispersion direction is along a column of pixels.
For comparison lamp observations, the central 100 pixel columns are extracted
and summed along the spatial axis, weighted by the inverse-variances of each
pixel.
Source spectral traces were always placed in this region during observations,
and the curvature of the wavelength solution over this region is negligible.
For source observations, 1D spectra are extracted using an empirical
estimate of the pixel-convolved line spread function (LSF): at each of 16 pixel
rows between indices 750 and 850 (i.e. around the effective center of the
dispersion axis), a model for the LSF is fit to the 1D trace.
The LSF is assumed to be a Voigt profile with parameters $A$, amplitude; $x_0$,
profile centroid; $\sigma_G$, Gaussian root-variance; $\gamma_L$, Lorentzian
half-width at half maximum, and the background is modeled using a linear
polynomial with parameters $\alpha_1$ and $\alpha_2$, the slope and intercept
at the profile centroid, respectively.
The maximum-likelihood LSF + background model parameters are estimated from
each pixel row using \lbfgsb\ optimization (\citealt{Zhu:1994}), implemented in
the \package{scipy} package.
From the 16 fits, the median profile width parameters, $\sigma_G$ and
$\gamma_L$, are then taken to be the LSF width parameters.
At each of 1600 rows in each source spectrum image, the LSF amplitude and
centroid, and background slope and intercept at line centroid, are then fit
using the same procedure as above but with the width parameters fixed.
The LSF amplitudes are stored as the source fluxes, and the background model
intercepts are stored as the background (sky) fluxes.
Occasionally, nearby sources fall in the slit; these are masked by hand in the
2D images and do not bias the LSF extraction fits.

@@

This wavelength calibration procedure produces a nightly model for the
nonlinear dispersion of wavelength as a function of pixel value for each
extracted 1D spectrum.
However, because of flexure in the spectrograph, each observation will have an
additional offset of, typically, $\approx 0.5$--$2~{\rm \AA}$, corresponding to
a significant velocity offset at \Ha\ of $\approx 25$--$100~\kms$ (discussed
below).
\figurename~\ref{fig:sample-spectra} shows a random sample of 10
wavelength-calibrated source spectra of comoving star pair members normalized
to 1 at $\lambda = 5500~{\rm \AA}$ and shifted by an arbitrary offset for
visualization.
As suggested by the features in these 10 spectra, there is a mix of
metallicities and spectral types amongst the observed targets.
Note that the spectra are not flux-calibrated, as we only care about measuring
radial velocities from this sample.
